---
output:
  html_document: default
  pdf_document: default
---


Visualize experiment:

```{r}
library(rgl)

start_cyl <- cylinder3d(cbind(0, 0, seq(0, 10, length = 10)), radius = c(3,3,3), sides = 20, closed = -2)
target_cyl <- cylinder3d(cbind(60, 0, seq(0, 10, length = 10)), radius = c(3,3,3), sides = 20, closed = -2)
cyl1 <- cylinder3d(cbind(0, 0, 10 + seq(0, 12.5, length = 10)), radius = c(3,3,3), sides = 20, closed = -2)
cyl2 <- cylinder3d(cbind(60, 0, 10 + seq(0, 12.5, length = 10)), radius = c(3,3,3), sides = 20, closed = -2)
cyl3 <- cylinder3d(cbind(30, 0, seq(0, 20, length = 10)), radius = c(3,3,3), sides = 10, closed = -2)
shade3d(addNormals(subdivision3d(start_cyl)), col = "darkgreen")
shade3d(addNormals(subdivision3d(target_cyl)), col = "darkgreen")
shade3d(addNormals(subdivision3d(cyl1)), col = "pink")
shade3d(addNormals(subdivision3d(cyl2)), col = "pink", alpha = 0.5)
shade3d(addNormals(subdivision3d(cyl3)), col = "lightblue")
surface3d(c(-7, 67), c(-20, 20), matrix(0, 2, 2), col = "black", alpha = 0.9, specular = "black")


colors <- c("blue", "red","yellow", "orange", "purple", "cyan", "brown", "green", "magenta", "salmon")

for (j in 1:10){
    for (i in 1:10){
    lines3d(armdata[[7]][[j]][[i]], col = colors[j])
  }
}
```


#1. Assess if there is a significant effect of person on the observed trajectories.

• First layer: Experiments.
• Second layer: Persons. 
• Third layer: Repetitions.

Format data into DF

```{r}
data <- as.data.frame(do.call(cbind, armdata))

data_col <- c()
for (person in 1:10){
  for (setup in 1:16){
    for (repetition in 1:10){
      data_col <- rbind(data_col,data[[setup]][[person]][[repetition]])
    }
  }
}

times <- rep(1:100,1600)

repetitions <- c()
for (k in 1:10){
  for (j in 1:16){
    for (i in 1:10){
      repetitions <- c(repetitions,rep(i, 100))
    }
  }
}
repetitions <- as.factor(repetitions)

setups <- c()
for (j in 1:10){
  for (i in 1:16){
    setups <- c(setups,rep(i,1000))
  }
}
setups <- as.factor(setups)

persons <- c()
for (j in 1: 10){
  persons <- c(persons, rep(j,16000))
}
persons <- as.factor(persons)

df <- data.frame(x = data_col[,1], y = data_col[,2], z = data_col[,3], time = as.factor(times),repetition = repetitions, setup = setups, person = persons)

length(df$x)
#df <- na.omit(df)
# length(df$x)




shitrows <- c(132001,134101,137001, 137002, 138001, 138002, 140001, 140002, 140003, 140004, 141101, 141102)
df[shitrows,]$x <- df[shitrows+1,]$x
df[shitrows,]$y <- df[shitrows+1,]$y
df[shitrows,]$z <- df[shitrows+1,]$z

shitrows <- c(137001, 138001, 140001, 140002, 140003, 141101)
df[shitrows,]$x <- df[shitrows+1,]$x
df[shitrows,]$y <- df[shitrows+1,]$y
df[shitrows,]$z <- df[shitrows+1,]$z

shitrows <- c(140001, 140002)
df[shitrows,]$x <- df[shitrows+1,]$x
df[shitrows,]$y <- df[shitrows+1,]$y
df[shitrows,]$z <- df[shitrows+1,]$z

shitrows <- c(140001)
df[shitrows,]$x <- df[shitrows+1,]$x
df[shitrows,]$y <- df[shitrows+1,]$y
df[shitrows,]$z <- df[shitrows+1,]$z

head(df)
length(df$x)
```
```{r}
#boxplot til z
boxplot(df$z ~ df$setup, 
        main = "", xlab="Setup",ylab="Z-coordinate",
        cex.main = 1.8, cex.lab = 1.2,
        col=brewer.pal(n = 10, name = "Dark2"))
```
```{r}

```


Fjerner outlier
```{r}
#dfoutlier = df[(df$person == 2) & (df$repetition != 7) & (df$setup == 5),]
#install.packages("RColorBrewer")
library("RColorBrewer")

for (i in 1:100){
  df[(df$person == 2) & (df$repetition == 7) & (df$setup == 5) & (df$time ==i),]$x = mean(df[(df$person == 2) & (df$repetition != 7) & (df$setup == 5) & (df$time ==i),]$x)
  
  df[(df$person == 2) & (df$repetition == 7) & (df$setup == 5) & (df$time ==i),]$y = mean(df[(df$person == 2) & (df$repetition != 7) & (df$setup == 5) & (df$time ==i),]$y)
  
  df[(df$person == 2) & (df$repetition == 7) & (df$setup == 5) & (df$time ==i),]$z = mean(df[(df$person == 2) & (df$repetition != 7) & (df$setup == 5) & (df$time ==i),]$z)
}


boxplot(df$z ~ df$person, 
        #main = "", xlab="Person",ylab="Z-coordinate",
        #cex.main = 1.8, cex.lab = 1.2,
        #col=brewer.pal(n = 10, name = "Dark2"))

```

Save to excel
```{r}
#write.csv(df,"armdatadf.csv", row.names = T)
```

Data exploration / description
```{r}

boxplot(df$y ~ df$person, 
        main = "Y-coordinate as a function of person", xlab="Person",ylab="Y-coordinate",
        cex.main = 2, cex.lab = 2,
        col=brewer.pal(n = 10, name = "Dark2"))
boxplot(df$y ~ df$setup)

boxplot(df$x, df$y, df$z)

summary(df$z)
summary(df$y)
summary(df$x)


```
Boxplot X-Y-Z coordinater med person
```{r}

boxplot(df$x ~ df$person, 
        main = "", xlab="Person",ylab="X-coordinate",
        cex.main = 1.8, cex.lab = 1.2,
        col=brewer.pal(n = 10, name = "Dark2"))
boxplot(df$y ~ df$person, 
        main = "", xlab="Person",ylab="Y-coordinate",
        cex.main = 1.8, cex.lab = 1.2,
        col=brewer.pal(n = 10, name = "Dark2"))
boxplot(df$z ~ df$person, 
        main = "", xlab="Person",ylab="Z-coordinate",
        cex.main = 1.8, cex.lab = 1.2,
        col=brewer.pal(n = 10, name = "Dark2"))
```

lave boxplot med setup istedet for persons - stadig med x y og z
```{r}
boxplot(df$x ~ df$setup, 
        main = "", xlab="Setup",ylab="X-coordinate",
        cex.main = 1.8, cex.lab = 1.2,
        col=brewer.pal(n = 10, name = "Dark2"))
boxplot(df$y ~ df$setup, 
        main = "", xlab="Setup",ylab="Y-coordinate",
        cex.main = 1.8, cex.lab = 1.2,
        col=brewer.pal(n = 10, name = "Dark2"))
boxplot(df$z ~ df$setup, 
        main = "", xlab="Setup",ylab="Z-coordinate",
        cex.main = 1.8, cex.lab = 1.2,
        col=brewer.pal(n = 10, name = "Dark2"))
```
En helt anden funktion med xyz, sqrt(dfx^2+dfy^2+dfz^2)
Vurdere længde udfra person og setup

```{r}
#tilføjer til dataframe
#normalisere x,y,z. 
#scale(x,center=True,scale=True)

df$len = sqrt(scale(df$x,center = TRUE, scale = TRUE)^2+scale(df$y,center = TRUE, scale = TRUE)^2+scale(df$z,center = TRUE, scale = TRUE)^2)

```

```{r}

boxplot(df$len ~ df$person,
        main = "", xlab="Person",ylab="Euclidean distance",
        cex.main = 1.8, cex.lab = 1.2,
        col=brewer.pal(n = 10, name = "Dark2"))
boxplot(df$len ~ df$setup,
        main = "", xlab="Setup",ylab="Euclidean distance",
        cex.main = 1.8, cex.lab = 1.2,
        col=brewer.pal(n = 10, name = "Dark2"))

```
P-værdier - euclid
```{r}
pvalseuclid <- rep(NA, 1600)

i = 1

for (t in 1:100){
  for (s in 1:16){
    df_local <- subset(df, time==t & setup == s)
    L <- lm(len ~ person, data = df_local)
    pvalseuclid[i] <- anova(L)$"Pr(>F)"[1]
    
    i = i+1
  }
}
```
p-adjusted - euclid
```{r}
pvalseuclid.adjusted <- p.adjust(pvalseuclid, method ="BH")

mean(pvalseuclid.adjusted<0.01)#1%
mean(pvalseuclid.adjusted<0.05)#5%
mean(pvalseuclid.adjusted<0.10)#10%

plot(sort(pvalseuclid.adjusted),xlab = "", ylab = "P-value", main= "P-values for significance of person, when predicting X-value")
abline(h = 0.05, col = "red")
legend("topleft", inset=0.1, fill=c("black","red"), legend = c("Sorted p-values", "0.05"))

hist(pvalseuclid.adjusted, breaks = 50)
abline(v = 0.05, col = "red")

```


Anovas
```{r}
pvalsx <- rep(NA, 1600)
pvalsy <- rep(NA, 1600)
pvalsz <- rep(NA, 1600)
i = 1

for (t in 1:100){
  for (s in 1:16){
    df_local <- subset(df, time==t & setup == s)
    L <- lm(x ~ person, data = df_local)
    pvalsx[i] <- anova(L)$"Pr(>F)"[1]
    
    L <- lm(y ~ person, data = df_local)
    pvalsy[i] <- anova(L)$"Pr(>F)"[1]
    
    L <- lm(z ~ person, data = df_local)
    pvalsz[i] <- anova(L)$"Pr(>F)"[1]
    i = i+1
  }
}

```

QQ plots for persons:
```{r}
for (i in 1:5){
  t <- sample(1:100,1)
  s <- sample(1:16,1)
  df_local <- subset(df, time==t & setup == s)
  Lx <- lm(x ~ person, data = df_local)
  Ly <- lm(y ~ person, data = df_local)
  Lz <- lm(z ~ person, data = df_local)
  
  plot(Lx, which = 2, main =sprintf("Time: %s, Setup: %s, X-coordinates",t,s),cex.main = 1)
  
  plot(Ly, which = 2, main =sprintf("Time: %s, Setup: %s",t,s),cex.main = 1)
  
  plot(Lz, which = 2, main =sprintf("Time: %s, Setup: %s",t,s),cex.main = 1)
}


```

QQ plots for setup
```{r}
for (i in 1:10){
  t <- sample(1:100,1)
  p <- sample(1:10,1)
  df_local <- subset(df, time==89 & person == 9)
  Lx <- lm(x ~ setup, data = df_local)
  Ly <- lm(y ~ setup, data = df_local)
  Lz <- lm(z ~ setup, data = df_local)
  
  plot(Lx, which = 2, main =sprintf("Time: %s, Person: %s,",t,p),cex.main = 1)
  
  plot(Ly, which = 2, main =sprintf("Time: %s, Person: %s,",t,p),cex.main = 1)
  
  plot(Lz, which = 2, main =sprintf("Time: %s, Person: %s,",t,p),cex.main = 1)
}
```
Residual plots
```{r}
for (i in 1:10){
  t <- sample(1:100,1)
  p <- sample(1:10,1)
  df_local <- subset(df, time==t & person == p)
  Lx <- lm(x ~ setup, data = df_local)
  Ly <- lm(y ~ setup, data = df_local)
  Lz <- lm(z ~ setup, data = df_local)
  
  plot(Lx, which = 1,main =sprintf("Time: %s, Person: %s,",t,p),cex.main = 1)
  
  plot(Ly, which = 1, main =sprintf("Time: %s, Person: %s,",t,p),cex.main = 1)
  
  plot(Lz, which = 1, main =sprintf("Time: %s, Person: %s,",t,p),cex.main = 1)
}
```


Adjusted p-values
```{r}
pvalsx.adjusted <- p.adjust(pvalsx, method ="BH")
pvalsy.adjusted <- p.adjust(pvalsy, method ="BH")
pvalsz.adjusted <- p.adjust(pvalsz, method ="BH")

mean(pvalsx.adjusted<0.01)#1%
mean(pvalsx.adjusted<0.05)#5%
mean(pvalsx.adjusted<0.10)#10%

mean(pvalsy.adjusted<0.01)#1%
mean(pvalsy.adjusted<0.05)#5%
mean(pvalsy.adjusted<0.10)#10%

mean(pvalsz.adjusted<0.01)#1%
mean(pvalsz.adjusted<0.05)#5%
mean(pvalsz.adjusted<0.10)#10%

plot(sort(pvalsz.adjusted),xlab = "", ylab = "P-value", main= "P-values for significance of person, when predicting Z-value")
abline(h = 0.05, col = "red")
legend("topleft", inset=0.1, fill=c("black","red"), legend = c("Sorted p-values", "0.05"))

plot(sort(pvalsy.adjusted),xlab = "", ylab = "P-value", main= "P-values for significance of person, when predicting Y-value")
abline(h = 0.05, col = "red")
legend("topleft", inset=0.1, fill=c("black","red"), legend = c("Sorted p-values", "0.05"))


plot(sort(pvalsx.adjusted),xlab = "", ylab = "P-value", main= "P-values for significance of person, when predicting X-value")
abline(h = 0.05, col = "red")
legend("topleft", inset=0.1, fill=c("black","red"), legend = c("Sorted p-values", "0.05"))

hist(pvalsx.adjusted, breaks = 50)
abline(v = 0.05, col = "red")

hist(pvalsy.adjusted, breaks = 50)
abline(v = 0.05, col = "red")

hist(pvalsz.adjusted, breaks = 50)
abline(v = 0.05, col = "red")
```


```{r}
pvalsx <- rep(NA, 1000)
pvalsy <- rep(NA, 1000)
pvalsz <- rep(NA, 1000)
i = 1

for (t in 1:100){
  for (p in 1:10){
    df_local <- subset(df, time==t & person == p)
    L <- lm(x ~ setup, data = df_local)
    pvalsx[i] <- anova(L)$"Pr(>F)"[1]
    
    L <- lm(y ~ setup, data = df_local)
    pvalsy[i] <- anova(L)$"Pr(>F)"[1]
    
    L <- lm(z ~ setup, data = df_local)
    pvalsz[i] <- anova(L)$"Pr(>F)"[1]
    i = i+1
  }
}

```


```{r}
pvalsx.adjusted <- p.adjust(pvalsx, method ="BH")
pvalsy.adjusted <- p.adjust(pvalsy, method ="BH")
pvalsz.adjusted <- p.adjust(pvalsz, method ="BH")

mean(pvalsx.adjusted<0.01)#1%
mean(pvalsx.adjusted<0.05)#5%
mean(pvalsx.adjusted<0.10)#10%

mean(pvalsy.adjusted<0.01)#1%
mean(pvalsy.adjusted<0.05)#5%
mean(pvalsy.adjusted<0.10)#10%

mean(pvalsz.adjusted<0.01)#1%
mean(pvalsz.adjusted<0.05)#5%
mean(pvalsz.adjusted<0.10)#10%

plot(sort(pvalsz.adjusted),xlab = "", ylab = "P-value", main= "P-values for significance of setup, when predicting Z-value")
abline(h = 0.05, col = "red")
legend("topleft", inset=0.1, fill=c("black","red"), legend = c("Sorted p-values", "0.05"))

plot(sort(pvalsy.adjusted),xlab = "", ylab = "P-value", main= "P-values for significance of setup, when predicting Y-value")
abline(h = 0.05, col = "red")
legend("topleft", inset=0.1, fill=c("black","red"), legend = c("Sorted p-values", "0.05"))


plot(sort(pvalsx.adjusted),xlab = "", ylab = "P-value", main= "P-values for significance of setup, when predicting X-value")
abline(h = 0.05, col = "red")
legend("topleft", inset=0.1, fill=c("black","red"), legend = c("Sorted p-values", "0.05"))

hist(pvalsx.adjusted, breaks = 50)
abline(v = 0.05, col = "red")

hist(pvalsy.adjusted, breaks = 50)
abline(v = 0.05, col = "red")

hist(pvalsz.adjusted, breaks = 50)
abline(v = 0.05, col = "red")
```






Interaction plots
```{r}
df_local <- subset(df)

interaction.plot(df_local$setup, df_local$person, df_local$y, col = c("blue", "red","yellow", "orange", "purple", "cyan", "brown", "green", "magenta", "salmon"), lty=1, main = "y as function of person / setup")

interaction.plot(df_local$setup, df_local$person, df_local$x, col = c("blue", "red","yellow", "orange", "purple", "cyan", "brown", "green", "magenta", "salmon"), lty=1, main = "x as function of person / setup")

interaction.plot(df_local$setup, df_local$person, df_local$z, col = c("blue", "red","yellow", "orange", "purple", "cyan", "brown", "green", "magenta", "salmon"), lty=1, main = "z as function of person / setup")

interaction.plot(df_local$person, df_local$setup, df_local$z, col = c("blue", "red","yellow", "orange", "purple", "cyan", "brown", "green", "magenta", "salmon"), lty=1, main = "z as function of setup / person")

interaction.plot(df_local$person, df_local$setup, df_local$y, col = c("blue", "red","yellow", "orange", "purple", "cyan", "brown", "green", "magenta", "salmon"), lty=1, main = "y as function of setup / person")

interaction.plot(df_local$person, df_local$setup, df_local$x, col = c("blue", "red","yellow", "orange", "purple", "cyan", "brown", "green", "magenta", "salmon"), lty=1, main = "x as function of setup / person")

```

